---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# yahtsee

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/yahtsee)](https://CRAN.R-project.org/package=yahtsee)
[![R-CMD-check](https://github.com/njtierney/yahtsee/workflows/R-CMD-check/badge.svg)](https://github.com/njtierney/yahtsee/actions)
[![Codecov test coverage](https://codecov.io/gh/njtierney/yahtsee/branch/main/graph/badge.svg)](https://codecov.io/gh/njtierney/yahtsee?branch=main)
<!-- badges: end -->

The goal of yahtsee is to provide tools around fitting hierarchical time series models for data commonly found in malaria.

## Installation

You can install the released version of yahtsee from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("njtierney/yahtsee")
```

### Installing INLA

This package requires `INLA` - if it didn't install, try the following to install it

```r
options(
  repos = c(
    INLA = "https://inla.r-inla-download.org/R/testing",
    CRAN = "https://cloud.r-project.org/"
    )
)

install.packages("INLA")
```

## Example

```{r example}
library(yahtsee)
# for nice table printing
library(tibble)
```


There are two example datasets.

1. `who_regions`, containing region information from WHO database

```{r who-regions-data}
who_regions
```

2. `malaria_africa_ts` - containing malaria prevalence data, extracted using the `malariaAtlas` R package - https://cran.r-project.org/web/packages/malariaAtlas/vignettes/overview.html

```{r malaria-africa-ts}
malaria_africa_ts
```

